DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-09-29 paper/intro.tex

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies for ever more autonomous decision making
capabilities---ultimately fully autonomous driving capability without
any human intervention. Deep neural networks (DNNs) have been
sucessfully applied in various perception and control tasks in 
recent years, and they are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized SoC
(MobileEye EyeQ), which uses a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNNs based Artifical Intelligence workloads may be used in future
autonomous vehicles.

% need testbeds.
% but challenges 
%   (1) real-time processing (2) cost (3) safety (4) SWaP
Developing and testing DNN models for autonomous vehicles is
however, challenging for a variety of reasons. The first challegen is
finding a suitable computing platform. It is generally known that DNN
models are computationally demainding. Furtheremore, because many of
the AI workloads, such as models for end-to-end control or perception,
have strict real-time requirements as cars have to interact with the
physical world in real-time. On the other, the computing platform must
also satisfy cost, size, weight, and power constraints. 
The second challenge is preparing a
physical car platform to collect data and test the software. While
instrumenting a real-car would be most realistic, as many
companies such as google and uber are doing, but doing it is costly
and would pose safety concerns. 
% our goals:
%   (1) build a safe and low cost testbed but utilizing the
%       state-of-the-art AI technologies: DNN based end-to-end control
%   (2) analyze real-time issues in DNN based end-to-end control.
%   (3) evalute popular embedded platforms for the AI workloads
%       (end-to-end control, perception).
%       - Can apsberry pi handle AI workloads?
%       - Can embedded GPU do better? How much?
Therefore, there is a strong need for low-cost testbeds that still
employs state-of-the-art AI technologies and are capable of executing
real-time autonomous vehicle operations.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research and education. From hardware perspective, 
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$200 in total).
The DeepPicar, however, employs state-of-the-art AI
techonologies, including a vision-based end-to-end control system that
utilizes a deep convolutional nueral network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}. 
The network is \emph{deep} as it has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is in fact identifical to the one
used in NVIDIA's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
than the obvious difference in scale (RC car vs. real car) the only
differences between the two systems---from the computing
perspective---are that our system is implemented in 
TensorFlow~\cite{abadi2016tensorflow} and runs on a 
Raspberry Pi 3 whereas NVIDIA's DAVE-2 systems is implemented in Torch
7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
automotive specialized computing system~\cite{drivepx}), which is more
powerful but also more expensive.

% how we trained and tested (in terms of network performance)
We follow the standard supervised learning methodology to train the
system with a goal of following lanes on the ground, mimicking real
roads. We first collected data for taining and validation by manually 
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform interference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each \emph{inference} operation must
be completed within the desired control period. (e.g., 50ms period for
20Hz control frequency.)

% how we evaluated (in terms real-time performance)
% what are our findings?

@@

%% We chose to use a Raspberry Pi 3 not only because of its
%% affordability and availability but also because it is representative
%% of today's mainstream low-end multicore processors, found in
%% smartphones, tables and other embedded devices.
%% The DeepPicar is comprised of a Raspberry
%% Pi 3 Model B, a camera and a low cost RC car.

>>>>

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability without
any human intervention. Deep neural networks (DNNs) have been
sucessfully applied in various perception and control tasks in 
recent years, and they are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which uses a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based Artifical Intelligence workloads may be used in future
autonomous vehicles.

% big problem
Executing these AI workloads on a embedded computing platform that can
be used in a car poses several additional challenges. First, many AI
workloads in vehicles are computationaly demanding and have strict
real-time requirements. For example, latency in a vision-based object
detection task may directly linked to safety of the vehicle. This
requires a high computing capacity as well as means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require
highly efficient computing platform. These two conflicting
requirements  complicate the platform selection process as observed in
~\cite{Otterness2017}, which evaluated vision workloads on
NVIDIA's Tegra TX1 platform.
%% For example, while today's self-driving car
%% prototype equip more \$100,000 
%% of computers and sensors~\cite{juliussen2014emerging}, a study
%% found that aveage consumers are willing to pay much less amount of
%% extra cost for a self-driving capability~\cite{Daziano2017}.
%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/

% related work and remaing problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While a real car
testbed would be most ideal, it is not only highly expensive but also
poses serious safety concerns that hinder development and exploration.
Therefore, there is a strong need for safer and less costly testbeds
that still employ state-of-the-art AI technologies. There are already
several relatively inexpensive RC-car based testbeds, such as MIT
RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
However, these RC-car testbeds still cost more than \$3,000.

% our goals
Instead, we want to build a low cost testbed that still employs the
state-of-the art AI technologies. Specifically, we focus on a deep
convolutional neural network (CNN) based end-to-end control system,
which was developed for a real self-driving car, NVIDIA
DAVE-2~\cite{Bojarski2016}, and use the same moethodology on a
smaller, safer and less costly setup. In developing the testbed, our
goals are (1) analyze real-time issues in DNN based end-to-end
control; (2) evalute real-time performance of contemporarly embedded
multicore platforms for the AI workloads.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
platform for research and education. From hardware perspective, 
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$200 in total).
The DeepPicar, however, employs state-of-the-art AI
techonologies, including a vision-based end-to-end control system that
utilizes a deep convolutional nueral network (CNN).
The network receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}. 
The network has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
The network architecture is identifical to that of NVIDIA's DAVE-2
self-driving car~\cite{Bojarski2016}, which uses much more powerful
computer (NVIDIA's automotive specialized DRIV-PX
computer~\cite{drivepx}) than the Raspberry Pi~3.
We chose to use the Pi 3 not only because of its
affordability but also because it is representative
of today's mainstream low-end multicore platforms, found in
smartphones and other embedded devices.

%% Other than the difference in scale (RC car vs. real car), the only other
%% differences between the two systems---from the computing
%% perspective---are that our system is implemented in
%% TensorFlow~\cite{abadi2016tensorflow} and runs on a Raspberry Pi 3
%% whereas NVIDIA's DAVE-2 systems is implemented in Torch
%% 7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
%% automotive specialized computing system~\cite{drivepx}), which is more
%% powerful but also more expensive.

% how we trained (maybe moved to a later section)
We follow the standard supervised learning methodology to train the
system to follow tracks on the ground. We collect data for
taining and validation by manually  
controlling the RC car and recording the vision (from the webcam
mounted on the RC-car) and the human control inputs. We then train the
network offline using the collected data on a desktop computer, which
equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
back to the Raspberry Pi, which is then used to perform interference
operations---locally on the Pi---in the car's main control loop in
real-time. For real-time control, each inference operation must
be completed within the desired control period. (e.g., 50ms period for
20Hz control frequency.)
% how we evaluated (in terms real-time performance)

% what are our findings?
\fixme{TBD: Summary of our findings}

@@

%% The DeepPicar is comprised of a Raspberry
%% Pi 3 Model B, a camera and a low cost RC car.
